\pkg{webchem}: An \proglang{R} Package to Retrieve Chemical Information from the Web

### 2020-03-29 revision3_2581/paper.tex

\author{Eduard Sz\"ocs\\Universit\"at Koblenz-Landau \And
	      Tam\'as Stirling\\Independent Developer \And
	      Andreas Scharm\"uller\\Universit\"at Koblenz-Landau \And
        Ralf B. Sch\"afer\\Universit\"at Koblenz-Landau}
\title{\pkg{webchem}: An \proglang{R} Package to Retrieve Chemical Information from the Web}

\Plainauthor{Eduard Sz\"ocs, Tam\'as Stirling, Andreas Scharm\"uller, Ralf B. Sch\"afer}

@@

We use these data to retrieve basic information on the substances.
%
\begin{CodeChunk}
\begin{CodeInput}
R> etox_data <- etox_basic(ids$etoxid)
\end{CodeInput}
\end{CodeChunk}
%
When possible, \pkg{webchem} returns a data frame with one row per substance.  However, data from some sources can be very voluminous and not tabular. In these cases \pkg{webchem} always returns a named list (one entry for each substance).
Therefore, we provide extractor functions for the common identifiers: CAS, SMILES and InChIKeys.
%
\begin{CodeChunk}
\begin{CodeInput}
R> etox_cas <- cas(etox_data)
R> head(etox_cas)
\end{CodeInput}
\begin{CodeOutput}
       8668        8494        <NA>        8397        7240        7331 
 "105-67-9" "1570-64-5"          NA "1912-24-9"   "71-43-2" "6190-65-4" 
\end{CodeOutput}
\end{CodeChunk}
%
A variety of data are available and we cannot provide extractor
functions for each of those.  Therefore, if users need to extract
other data, they have to write simple extractor functions (see the
following examples).

@@

\begin{CodeChunk}
\begin{CodeInput}
R> cas_rns <- lc50[order(lc50$value)[1:3],"cas"]
R> chebiids <- get_chebiid(cas_rns)
R> chebiids <- do.call(rbind, chebiids)
R> comp <- chebi_comp_entity(chebiids$chebiid)
R> pars <- lapply(comp, function(x) with(x, parents[parents$type == "has role",]))
\end{CodeInput}
\end{CodeChunk}

Finally we can compile the data into a table. Table~\ref{tab:chebi} provides information on the individual roles of the three most toxic chemicals and we observe that, for example ethion, the most toxic pesticide towards \textit{D. magna} in our table, is designed to be used as an acari- and insecticide, which inhibts the acetylcholinesterase enzyme.

@@

\subsection[Querying partitioning coefficients]{Querying partitioning coefficients}
Some data sources also provide data on chemical properties that can be
queried.  Here we query for the \code{lc50} data predicted
$\mathrm{log}~P_{oct/wat}$ from the PubChem database to build a
simple quantitative structure-activity relationship (QSAR) to predict
toxicity. The resulting data and model are displayed in Figure~\ref{fig:fig3}.
%
\begin{CodeChunk}
\begin{CodeInput}
R> cid <- get_cid(lc50$cas)
R> pc_data <- pc_prop(cid)
R> lc50$logp <- pc_data$XLogP
\end{CodeInput}
\end{CodeChunk}
%

>>>>

\author{
	Eduard Sz\"ocs\\Universit\"at Koblenz-Landau \And
	Tam\'as Stirling\\Independent Developer \And
	Eric Scott\\Tufts University \And
	Andreas Scharm\"uller\\Universit\"at Koblenz-Landau \And
    Ralf B. Sch\"afer\\Universit\"at Koblenz-Landau}
\title{\pkg{webchem}: An \proglang{R} Package to Retrieve Chemical Information from the Web}

\Plainauthor{Eduard Sz\"ocs, Tam\'as Stirling, Eric Scott, Andreas Scharm\"uller, Ralf B. Sch\"afer}

@@

We use these data to retrieve basic information on the substances.
%
\begin{CodeChunk}
\begin{CodeInput}
R> etox_data <- etox_basic(ids$etoxid)
\end{CodeInput}
\end{CodeChunk}
%
When possible, \pkg{webchem} returns a data frame with one or more rows per substance.  However, data from some sources can be very voluminous and not tabular. In these cases \pkg{webchem} always returns a named list (one entry for each substance).
We provide extractor functions for the common identifiers: CAS, SMILES and InChIKeys.
%
\begin{CodeChunk}
\begin{CodeInput}
R> etox_cas <- cas(etox_data)
R> head(etox_cas)
\end{CodeInput}
\begin{CodeOutput}
       8668        8494        <NA>        8397        7240        7331 
 "105-67-9" "1570-64-5"          NA "1912-24-9"   "71-43-2" "6190-65-4" 
\end{CodeOutput}
\end{CodeChunk}
%
A variety of data are available and we cannot provide extractor
functions for each of those.  Therefore, if users need to extract
other data, they have to write simple extractor functions (see the
following examples).

@@

\begin{CodeChunk}
\begin{CodeInput}
R> cas_rns <- lc50[order(lc50$value)[1:3],"cas"]
R> chebiids <- get_chebiid(cas_rns)
R> comp <- chebi_comp_entity(chebiids$chebiid)
R> pars <- lapply(comp, function(x) with(x, parents[parents$type == "has role",]))
\end{CodeInput}
\end{CodeChunk}

Finally we can compile the data into a table. Table~\ref{tab:chebi} provides information on the individual roles of the three most toxic chemicals and we observe that, for example ethion, the most toxic pesticide towards \textit{D. magna} in our table, is designed to be used as an acari- and insecticide, which inhibits the acetylcholinesterase enzyme.

@@

\subsection[Querying partitioning coefficients]{Querying partitioning coefficients}
Some data sources also provide data on chemical properties that can be
queried.  Here we query for the \code{lc50} data predicted
$\mathrm{log}~P_{oct/wat}$ from the PubChem database to build a
simple quantitative structure-activity relationship (QSAR) to predict
toxicity. The resulting data and model are displayed in Figure~\ref{fig:fig3}.
%
\begin{CodeChunk}
\begin{CodeInput}
R> cid <- get_cid(lc50$cas, first = TRUE)
R> pc_data <- pc_prop(cid)
R> lc50$logp <- pc_data$XLogP
\end{CodeInput}
\end{CodeChunk}
%
